Temporal graph realization from fastest paths

### 2024-01-19 SAND2024/SAND24-longversion.tex
Updates from Overleaf

Remember, in the case where the clip vertex $u$ of the tree $T_u$ in $G[Z \cup \{u\}]$ is a vertex of interest, we split the vertices in the first layer of $T_u$ into at most $|N_{G'}(u)|$ equivalence classes (as explained in \cref{sec:preprocessing-FPT}).
Let us now show the following important property of these equivalence classes.
\begin{lemma}\label{lemma:treeEquivClass-intVertices}
    Let $u \in V(G')$ be a clip vertex of the tree $T_u$ in $G[Z \cup \{u\}]$, where $u \in U$,
    and let $z_1,z_2 \in V(T_u)$ be in the same equivalence class of the tree $T_u$.
    Then, the fastest temporal paths from $z_1$ and from $z_2$ to any other vertex in $G'$ coincide on the edges in $G'$.
    Similarly, the fastest temporal paths from any other vertex in $G'$ to $z_1$ and to $z_2$ coincide on the edges in $G'$.
\end{lemma}
%
\begin{proof}
    Let $y \neq u$ be a vertex in $V(G')$.
    %
    Denote with $P_1$ the underlying path of the fastest temporal path from $z_1$ to $y$, which consists of the edge $z_1 u$ and the path $P$ from $u$ to $y$.
    Similarly, let $Q_2$ be the underlying path of the fastest temporal path from $z_2$ to $y$, consisting of the edge $z_2u$ together with the path $Q$ from $u$ to $y$.
    Define $P_2$ as the second path from $z_2$ to $y$ that first uses the edge $z_1u$ and then the path $P$.
    Similarly, $Q_1$ represents the second path from $z_1$ to $y$ that first uses the edge $z_2u$ and then the path $Q$.
    Our objective is to demonstrate that either $P = Q$ or that $d(P_1, \lambda) = d(Q_1, \lambda)$ (and $d(P_2, \lambda) = d(Q_2, \lambda)$).
    This implies that $z_1$ and $z_2$ use temporal paths that coincide on the vertices of $V(G) \setminus V(T_u)$ to reach $y$.

>>>>

Remember, in the case where the clip vertex $u$ of the tree $T_u$ in $G[Z \cup \{u\}]$ is a vertex of interest, we split the vertices in the first layer of $T_u$ into at most $|N_{G'}(u)|$ equivalence classes (as explained in \cref{sec:preprocessing-FPT}).
Let us now show the following important property of these equivalence classes.
\begin{lemma}\label{lemma:treeEquivClass-intVertices}
    Let $u \in V(G')$ be a clip vertex of the tree $T_u$ in $G[Z \cup \{u\}]$, where $u \in U$,
    and let $z_1,z_2 \in V(T_u)$ be in the same equivalence class of the tree $T_u$.
    Then, the fastest temporal paths from $z_1$ and from $z_2$ to any other vertex in $G'$ coincide on the edges in $G'$.
    Similarly, the fastest temporal paths from any other vertex in $G'$ to $z_1$ and to $z_2$ coincide on the edges in $G'$.
\end{lemma}
%
\begin{proof}
    Let $y \neq u$ be a vertex in $V(G')$.
    %
    Denote with $P_1$ the underlying path of the fastest temporal path from $z_1$ to $y$, which consists of the edge $z_1 u$ and the path $P$ from $u$ to $y$.
    Similarly, let $Q_2$ be the underlying path of the fastest temporal path from $z_2$ to $y$, consisting of the edge $z_2u$ together with the path $Q$ from $u$ to $y$.
    Define $P_2$ as the second path from $z_2$ to $y$ that first uses the edge $z_1u$ and then the path $P$.
    Similarly, $Q_1$ represents the second path from $z_1$ to $y$ that first uses the edge $z_2u$ and then the path $Q$.
    Our objective is to demonstrate that either $P = Q$ or that $d(P_1, \lambda) = d(Q_1, \lambda)$ (and $d(P_2, \lambda) = d(Q_2, \lambda)$).
    This implies that $z_1$ and $z_2$ use temporal paths that coincide on the vertices of $V(G) \setminus V(T_u)$ to reach $y$. For an illustration see \cref{fig:FPT-equivClass}.

    \begin{figure}[t]
	\centering
	\includegraphics[width=0.6\columnwidth]{fig-fpt-equivClass}
	\caption{An example of the situation in \cref{lemma:treeEquivClass-intVertices}.
        We have two paths $Q$ (in blue) and $P$ (in red) from $x$ to $y$.
        We assume that the fastest temporal path from $z_1$ to $y$ is $P_1 = \{z_1 u\} \cup P$,
		and the fastest temporal path from $z_2$ to $y$ is $Q_2 = \{z_2 x\} \cup Q$.
        %
        We denote with $Q_1 = \{z_1 u\} \cup Q$ another temporal path from $z_1$ to $y$, 
        with $P_2 = \{z_2 u\} \cup P$ another temporal path from $z_2$ to $y$.
        The labels on the edges $z_1 x, z_2 x$ are $t_1,t_2$, respectively. Similar, the labels of the last edges of paths $P$ and $Q$ are  $t_p, t_q$, respectively.
		\label{fig:FPT-equivClass}}
\end{figure}
